---
title: "C5_Assignment1"
author: "Mohsin"
date: "March 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


library(ggplot2)
library(knitr) 
library(markdown)
library(Hmisc)
library(lubridate)




# (1) Total Steps Taken in a Day
```{r, echo=TRUE}

datac<-read.csv("./Course5/activity.csv") 
str(datac)
names(datac)
steps <- aggregate(datac$steps, by = list(datac$date), sum, na.rm=TRUE) 
hist(steps[,2],main="Histogram:Total Steps Taken in a Day",col="pink", border="red",xlab="Step Count")

```


# (2) Calculate Mean and Median of the Total Number of Steps Taken Daily
```{r,echo=TRUE}
steps_mean <- mean(steps[,2]) 
steps_median <- median(steps[,2])
print(steps_mean)
print(steps_median)
```

# (3) Average Daily Pattern of Activity
```{r,echo=TRUE}
stepavg <- aggregate(steps ~ interval, data = datac, mean, na.rm = TRUE)

plot(stepavg$interval, stepavg$steps, type = "l", lwd = 2.5, col = "blue",
     main = "Average/Mean of Steps Taken", xlab = "Five Minute interval", ylab = "Steps(average)")
```


# (4) Inputting Missing Values
## This task is achieved by the 'Hmisc' package function called impute which fills in the missing values with arbitrary constants: You can search to fetch the help contents '?impute'
```{r,echo=TRUE}
library(Hmisc)
missingval <- length(which(is.na(datac$steps)))
imputed <- datac
imputed$steps <- impute(datac$steps, fun=mean)
imputedstepsd <- tapply(imputed$steps, imputed$date, sum)
hist(imputedstepsd,main="Histogram of Total Number of Steps(Missing Values Imputed)",col="beige",border = "brown",xlab="Imputed_Steps")
```
 

#(5) Weekday and Weekend Steps

```{r, echo=TRUE}
library(ggplot2)
ds<-datac$date
dss<-as.Date(ds)
dayz<-weekdays(dss)
datac$dayz<-dayz
dayz<-as.character(dayz)
wkday<-subset(datac,dayz==c("Monday","Tuesday","Wednesday","Thursday","Friday"))
wkend<-subset(datac,dayz==c("Satuday","Sunday"))
par(mfrow=c(1,2))
wkday_sb<-aggregate(steps~interval,wkday,mean,na.rm=TRUE)
wkend_sb<-aggregate(steps~interval,wkend,mean,na.rm=TRUE)  
plot1<- ggplot(wkday_sb, aes(interval, steps))
plot1 + geom_line(color = "pink",size=1)+theme_bw()+ggtitle("Weekday")
plot2<- ggplot(wkend_sb, aes(interval, steps))
plot2 + geom_line(color = "brown",size=1)+theme_bw()+ggtitle("Weekend")
```